---
title: "Tufte Stufte"
excerpt: "Sidenotes are pretty great. Also, I made a Bezier curve."
author: "Cameron Pfiffer"
date: '2018-02-26'
categories: []
tags: []
---

```{r setup, include=FALSE}
library(tufte)
library(ggplot2)
library(highcharter)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

This is _not_ a margin note.`r tufte::margin_note("This _is_ a margin note. As you can see, it is to the right of the left, and it indeed appears to be smaller.")` Were it a margin note as one might presuppose, it would be to the right of the left and perhaps a bit smaller. 

`r tufte::newthought("In the having of a new thought,")` I find myself overcome with thoughtfulness. You might say, perhaps you would be! `r tufte::margin_note("You would be overcome with thoughtfulness, as it is a thought.")`

Thank you for saying so. 

Quick, look at some math! `r tufte::margin_note("$$ x = 2+2 $$")` Tell me, what is $x$? 

Even better, here is a chart with made-up numbers. Look how fancy and _**interactive**_ it is.

```{r echo=FALSE}
n <- 15
dta <- dplyr::data_frame(
  x = rnorm(n),
  y = 1.5 * x + rnorm(n))

highchart() %>%
  hc_chart(type = "scatter") %>% 
  hc_add_series(data = list_parse(dta)) %>% 
  hc_title(text = "Some Random Numbers") %>% 
  hc_add_theme(hc_theme_tufte())
```

`r tufte::margin_note("Don't worry, it's quite boring -- you're not missing out on anything.")`But no, in all seriousness, I thought I'd take the chance to explore [Bezier curves](https://en.wikipedia.org/wiki/B%C3%A9zier_curve) while I'm mucking about with some fun looking theming. Bezier curves came up at work, the details of which I don't really know whether or not I can share. So, I won't.

```{r}
bezier <- function(t, p) {
  # A simple Bezier curve
  x = (1 - t) * (1 - t) * p[1, 1] + 2 * (1 - t) * t * p[2,1] + t * t * p[3,1]
  y = (1 - t) * (1 - t) * p[1, 2] + 2 * (1 - t) * t * p[2,2] + t * t * p[3,2]
  
  return(c(x,y))
}

points <- matrix(c(0, 5, 1, 0, 2, 3), ncol = 2)
line <- matrix(bezier(0:100/100, points), ncol=2)
head(line)
```
`r tufte::margin_note("Basically, we just plot a bunch of smaller points on a new, continuously defined function that has a nifty little bend in it.")`

```{r echo=FALSE}
highchart() %>%   
  hc_chart(type = "line") %>% 
  hc_add_series(data = line) %>% 
  hc_title(text = "A Bezier Curve") %>% 
  hc_add_theme(hc_theme_tufte())
```

Neato. What a great Bezier curve. Thanks to [xan](https://stackoverflow.com/a/5634528) on StackOverflow for providing the pseudocode for that Bezier curve.

Now, I'd like to ask you to hang out with that curve for a little while and consider it's elegance, if you would be so kind.